A tutorial on the model-dependent analysis of neutron and X-ray reflectometry

### 2020-07-30 refl_maths/paper.tex
add tikz image
fix formating

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage[separate-uncertainty=true]{siunitx}
\usepackage{mhchem}

@@

\section{Recursive methods}
The breakdown of the kinematic approach leads to the application of the Abel\`{e}s \cite{abeles_sur_1948} or Parratt \cite{parratt_surface_1954}, recursive model for the reflection of light at a given number of stratified interfaces, also referred to as the dynamical approach. 
This is an alternative methods for calculating the reflected intensity from a model, however, now the model is a series of layers of different scattering length density. 
The probing radiation can be either reflected or refracted at the interface between each of the layers. 
%
\begin{figure}[t]
    \includegraphics[width=0.49\textwidth]{refr}
    \caption{A schematic diagram showing the reflected ($r$) and transmitted ($t$) waves when an incident ($i$) wave enters an interface of thickness $d$.}
    \label{fig:refr}
\end{figure}
% 
Figure~\ref{fig:refr} shows the refraction and reflections from a simple two layer system, where layer \num{0} is a vacuum above the sample. 
It is clear, by analogy to Bragg's law, how the two waves, labelled $r$, could interfere constructively or destructively depending on the thickness of layer \num{1}, $d$.

@@

Figure \ref{fig:phase_problem_sld} shows two hypothetical samples of the same total thickness,
consisting of one and two layers, deposited on a substrate with
$\rho_s = \scinot{2.07}{-6} \angstrom^{-2}$. Even though the SLD depth profiles are distinctly
different, the resultant reflectivity curves are virtually identical,
with only a slight discrepancy at the first minimum. This ambiguity,
known as \textit{the phase problem}, is
an example of the information lost in diffraction experiments due
to the inability of measuring the phase of the scattered wave
--Notice that equation \ref{equ:kine} involves a modulus-squared.

\begin{figure}
    \includegraphics[width=0.49\textwidth]{phase_problem_SLD.pdf} 
    \caption{Two different samples $\rho_{1,2}(z)$ deposited on the same substrate $\rho_s$ (top panel)
    are indistinguishable from one another by looking only at their reflectivity curve $R(q)$
    (bottom panel).}
    \label{fig:phase_problem_sld}
\end{figure}

An heuristic explanation for the phase problem
can be built in terms of the auto-correlation function of $\rmd\rho/\rmd z$, which gives a
real-space representation of the information contained in the reflectivity pattern:

\begin{equation}
    \begin{split}
        \acf_{\rho'}(z) & = \int_{-\infty}^{\infty} \rho'(z)^* \, \rho'(z + \xi) \, \rmd \xi \\
                        & = \int_{-\infty}^{\infty} \left| \widehat{\rho'}(q)  \right|^{2} \, \rm{e}^{i\,q\,z} \, \rmd q,
    \end{split}
    \label{equ:acf}
\end{equation}

\noindent where the hat $\widehat{[]}$ indicates a Fourier transform, the star $[]^*$ a complex conjugate
and the prime $[]'$ a derivative.

\begin{figure}
    \includegraphics[width=0.49\textwidth]{phase_problem_ACF.pdf}
    \caption{Top panel: the derivatives of the two hypothetical SLD profiles discussed, $\rho_{1,2}'(z)$, are
    scaled $\delta$ functions peaking at the interfaces between SLD layers. Bottom panel: In contrast,
    the autocorrelation function, $\acf_{\rho'}(z)$ is almost identical for both samples,
    peaking at exactly the same locations.}
    \label{fig:phase_problem_acf}
\end{figure}

As $\rho'(z)$ is purely real and in both examples is only composed of 
$\delta$ functions, the ACF can be obtained by direct inspection
(See Figure \ref{fig:phase_problem_acf} and Equation \ref{equ:acf}):

$\bullet$ At zero lag, there is always full correlation, thus the ACF will have 
a large positive peak at the origin for both samples.

$\bullet$ For both samples, the interference between the front and back interfaces, at $z = 0$ and
$z = 200 \angstrom$ respectively, translates into a pair of $\delta$ peaks
with opposite signs in $\rho'$. Thus, at a 
$\xi = \pm 200 \angstrom$ lag, the ACF must peak with negative amplitude.

$\bullet$ For the two-layer sample, an extra pair of peaks in the ACF could be expected
at lags of $\xi = \pm 100 \angstrom$; however, since there is a complete cancellation between
the positive and the negative contributions of the $\delta$ functions at $z = 0$ and $z = 200$,
the expected extra pair of peaks is absent.

\noindent As the sld profiles of both samples translate into the same ACF,
it is now not surprising the similarity between their reflectivity curves.

@@

%%%%

>>>>

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage[separate-uncertainty=true]{siunitx}
\usepackage{mhchem}
\usepackage{tikz}
\usepackage{tikzscale}

@@

\section{Recursive methods}
The breakdown of the kinematic approach leads to the application of the Abel\`{e}s \cite{abeles_sur_1948} or Parratt \cite{parratt_surface_1954}, recursive model for the reflection of light at a given number of stratified interfaces, also referred to as the dynamical approach. 
This is an alternative methods for calculating the reflected intensity from a model, however, now the model is a series of layers of different scattering length density. 
The probing radiation can be either reflected or refracted at the interface between each of the layers. 
%
\begin{figure}[t]
    \includegraphics[width=\linewidth]{dynamic}
    \caption{A schematic diagram showing the reflected ($r$) and transmitted ($t$) waves when an incident ($i$) wave enters an interface of thickness $d$.}
    \label{fig:refr}
\end{figure}
%
Figure~\ref{fig:refr} shows the refraction and reflections from a simple two layer system, where layer \num{0} is a vacuum above the sample. 
It is clear, by analogy to Bragg's law, how the two waves, labelled $r$, could interfere constructively or destructively depending on the thickness of layer \num{1}, $d$.

@@

Figure \ref{fig:phase_problem_sld} shows two hypothetical samples of the same total thickness, consisting of one and two layers, deposited on a substrate with $\rho_s = \scinot{2.07}{-6} \angstrom^{-2}$. 
Even though the SLD depth profiles are distinctly different, the resultant reflectivity curves are virtually identical, with only a slight discrepancy at the first minimum. 
This ambiguity, known as \textit{the phase problem}, is an example of the information lost in diffraction experiments due to the inability of measuring the phase of the scattered wave (note that equation \ref{equ:kine} involves a modulus-squared).
\begin{figure}
    \includegraphics[width=0.49\textwidth]{phase_problem_SLD.pdf} 
    \caption{Two different samples $\rho_{1,2}(z)$ deposited on the same substrate $\rho_s$ (top panel)
    are indistinguishable from one another by looking only at their reflectivity curve $R(q)$
    (bottom panel).}
    \label{fig:phase_problem_sld}
\end{figure}

An heuristic explanation for the phase problem can be built in terms of the auto-correlation function of $\rmd\rho/\rmd z$, which gives a real-space representation of the information contained in the reflectivity pattern,
%
\begin{equation}
    \begin{split}
        \acf_{\rho'}(z) & = \int_{-\infty}^{\infty} \rho'(z)^* \, \rho'(z + \xi) \, \rmd \xi \\
                        & = \int_{-\infty}^{\infty} \left| \widehat{\rho'}(q)  \right|^{2} \, \rm{e}^{i\,q\,z} \, \rmd q,
    \end{split}
    \label{equ:acf}
\end{equation}
%
where the hat $\widehat{[]}$ indicates a Fourier transform, the star $[]^*$ a complex conjugate
and the prime $[]'$ a derivative. 
\begin{figure}
    \includegraphics[width=0.49\textwidth]{phase_problem_ACF.pdf}
    \caption{Top panel: the derivatives of the two hypothetical SLD profiles discussed, $\rho_{1,2}'(z)$, are
    scaled $\delta$ functions peaking at the interfaces between SLD layers. Bottom panel: In contrast,
    the autocorrelation function, $\acf_{\rho'}(z)$ is almost identical for both samples,
    peaking at exactly the same locations.}
    \label{fig:phase_problem_acf}
\end{figure}

As $\rho'(z)$ is purely real and in both examples is only composed of $\delta$ functions, the ACF can be obtained by direct inspection (See Figure~\ref{fig:phase_problem_acf} and Equation~\ref{equ:acf}): 
\begin{itemize}
    \item At zero lag, there is always full correlation, thus the ACF will have a large positive peak at the origin for both samples.
    \item For both samples, the interference between the front and back interfaces, at $z = 0$ and $z = 200 \angstrom$ respectively, translates into a pair of $\delta$ peaks with opposite signs in $\rho'$. Thus, at a $\xi = \pm 200 \angstrom$ lag, the ACF must peak with negative amplitude.
    \item For the two-layer sample, an extra pair of peaks in the ACF could be expected at lags of $\xi = \pm 100 \angstrom$; however, since there is a complete cancellation between the positive and the negative contributions of the $\delta$ functions at $z = 0$ and $z = 200$, the expected extra pair of peaks is absent.

As the sld profiles of both samples translate into the same ACF, it is now not surprising the similarity between their reflectivity curves.

@@
